Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2015-05-18 tex/scam/experiments.tex
a

\subsection{Simulations}
We first illustrate our methods using a simulation of the model
\begin{equation}\nonumber
         Y_i = f_0(\x_{iS}) + \epsilon_i \quad (i=1,2,\ldots,n).
\end{equation}
Here $\x_{i}$ denotes data sample $i$ drawn from some distribution
$P$, and $f_0$ is the true regression function. The variables
$\x_{iS}$ are a subset of $\x_i$ with dimension $|S|=s$, where $S$
represents the set of relevant variables, and $\epsilon_i$ is additive
noise drawn from $\mathcal{N}(0,\sigma^2)$. For all simulations, we set
$\sigma^2$ so that the signal-to-noise ratio(SNR,
$\frac{\trm{std}(Y)}{\sigma}$) is 5. Also, for all simulations except
the sixth, we choose the set of relevant variables $S$ uniformly at
random among all variables $\{1,...,p\}$.

@@

\subsubsection{First simulation}
In the \textrm{first simulation} (Figure~\ref{Support}a), we vary the
ambient dimension $p$. We set $Q$ to be one on the diagonal and $1/2$ on
the off-diagonal with $0.5$ probability, set $s=5$, and $p=64,128,256$
and $512$. We draw $X \sim N(0, I_p)$.  For each $(n,p)$ combination,
we generate $100$ independent trials.  In Figure~\ref{Support}(a), we
plot the probability of exact support recovery. We observe that the
algorithm performs consistent variable selection even if the
dimensionality is large. To give the reader a sense of the running
speed, for a dataset with $n=1,000$ and $p=512$, the code runs in
roughly two minutes on a machine with a 2.3 GHz Intel Core i5 CPU and 4
GB memory.

@@

\subsubsection{Third simulation}
In the \textrm{third simulation} (Figure~\ref{Support}d,e), we use the non-Gaussian distribution described above and set the covariance $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0, 0.2, 0.5, 0.9$. We use the non-additive $\Q$, same as in the first experiment, with $\alpha=0.5$ and fix $p=128, s=5$. We measure success not through exact recovery but through faithful recovery. We say that a trial is a successful if (1) all relevant variables were recovered and (2) fewer than $20$ variables were marked as relevant overall (true sparsity $s=5$). We use the same $\lambda$ as before. The probabilities of success are computed from 40 independent trials and plotted against various values of $\nu$ in Figure~\ref{Support}(d). Additionally, for $\nu = 0.5$, we show the number of selected variables versus the sample size as a box-and-whisker plot in Figure~\ref{Support}(e). As seen, for small to moderate correlations, AC/DC can successfully recover the relevant variables with only a small number of false positives.

In the fourth and fifth simulation, we use a softmax function as the ground truth
\begin{align}
f_0(x_{iS}) = \log \left( \sum_{k=1}^K \exp( \beta_k^\tran x_{iS} ) \right) - \mu
\end{align}
We generate random unit vectors as $\{\beta_k \in \R^s\}_{k=1,...,K}$ and choose $\mu$ so that $f_0$ has mean-zero. We set $K = 7$ for all the experiments.

\subsubsection{Fourth simulation}
For the \textrm{fourth simulation} (Figure~\ref{Support}f,g), we let $f_0$ be the softmax function and let $X$ be drawn from the boundary flat mixture distribution described earlier with the Toeplitz covariance $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0.5$. We set $s=5$ and vary $p=128,256,512$. We use the same faithful recovery criteria as the third simulation and plot the probability of faithful recovery against the number of samples in Figure~\ref{Support}(f). The probabilities are computed over 30 independent trials. Also, for $p=256$, we show the number of selected variables versus the sample size as a box-and-whisker plot in Figure~\ref{Support}(g). The softmax function is more challenging to estimate than the quadratic function; the softmax function requires about $n > 1500$ to achieve the same success probability as the quadratic function with $n=1000$. Regardless, we see that increasing the ambient dimension $p$ does not significantly affect the recovery probability.

@@

\subsubsection{Fifth simulation}
For the \textrm{fifth simulation} (Figure~\ref{fig:ac_v_dc}), we compare the variables selected via the AC stage and the variables selected via the DC stage. We use the softmax regression function and $X$ drawn from the boundary flat mixture distribution with a Toeplitz covariance and correlation level $\nu = -0.7$. We set $s=5, n=500, p=128$. We perform 30 independent trials and plot the frequency of variable selection in Figure~\ref{fig:ac_v_dc}. The true variables are $X_j$ for $j=5,6,7,8,9,10$. We plot the frequency of selection among only the first 20 variables, that is, $X_j$ for $j=1,...,20$. We do not plot selection frequencies for variables 21 to 128 because they are almost never selected by either AC or DC. As can be seen, the DC stage is slightly helpful in recovering the true variables but its effect is not significant. We thus believe that the DC stage, though important in theory, is not as important in practice; it may be omitted without significant detriment to the overall result.

@@

\begin{figure*}
\includegraphics[width=0.5\textwidth]{figs/ACvDC}
\caption{Frequency of variable selection among the first 20 variables ($X_j$ for $j=1,...,20$) in the AC stage vs. in the DC stage. The true variables are [5,6,7,8,9,10].}
\label{fig:ac_v_dc}
\end{figure*}

@@

We next use the Boston housing data rather than simulated data. This data set
contains 13 covariates, 506 samples and one response variable
indicating housing values in suburbs of Boston. The data and detailed description
can be found on the UCI Machine Learning Repository website\footnote{\url{http://archive.ics.uci.edu/ml/datasets/Housing}}.

@@

Next, in order to quantitatively study the predictive performance, we
run 3 times 5-fold cross validation, following the same procedure
described above---training, variable selection and refitting.  A plot
of the mean and standard deviation of the predictive mean squared
error (MSE) is shown in Figure \ref{Boston}(c). Since for AC/DC the same
regularization level $\lambda^{(t)}$ may lead to a slightly different number of selected
variables in different folds and runs, the values on the $x$-axis
for AC/DC are not necessarily integers. The figure clearly shows that AC/DC has a 
lower predictive MSE than LASSO.  We also compared the performance of
AC/DC with that of Additive Forward Regression (AFR) presented
in~\cite{Xi:09}, and found that they are similar.  The main advantages
of AC/DC compared with AFR and SpAM are that there are no smoothing
parameters required, and the optimization is formulated
as a convex program, guaranteeing a global optimum.

>>>>

\subsection{Simulations}
We first illustrate our methods using a simulation of the model
\begin{equation}\nonumber
         Y_i = f_0(\x_{iS}) + \epsilon_i \quad (i=1,2,\ldots,n).
\end{equation}
Here $\x_{i}$ denotes data sample $i$ drawn from some distribution
$P$, and $f_0$ is the true regression function. The variables
$\x_{iS}$ are a subset of $\x_i$ with dimension $|S|=s$, where $S$
represents the set of relevant variables, and $\epsilon_i$ is additive
noise drawn from $\mathcal{N}(0,\sigma^2)$. For all simulations, we set
$\sigma^2$ so that the signal-to-noise ratio(SNR,
$\frac{\trm{std}(Y)}{\sigma}$) is 5. Also, for all simulations except
the sixth, we choose the set of relevant variables $S$ uniformly at
random among all variables $\{1,\ldots,p\}$.

@@

\subsubsection{First simulation}
In the \textrm{first simulation} (Figure~\ref{Support}a), we vary the
ambient dimension $p$. We set $Q$ to be one on the diagonal and $1/2$ on
the off-diagonal with $0.5$ probability, set $s=5$, and $p=64,128,256$
and $512$. We draw $X \sim N(0, I_p)$.  For each $(n,p)$ combination,
we generate $100$ independent trials.  In Figure~\ref{Support}(a), we
plot the probability of exact support recovery. We observe that the
algorithm performs consistent variable selection even if the
dimensionality is large. To give the reader a sense of the running
speed, for a dataset with $n=1{,}000$ and $p=512$, the code runs in
roughly two minutes on a machine with a 2.3 GHz Intel Core i5 CPU and 4
GB memory.

@@

\subsubsection{Third simulation}
In the \textrm{third simulation} (Figure~\ref{Support}d,e), we use the
non-Gaussian distribution described above and set the covariance
to be $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0, 0.2, 0.5, 0.9$. We use the
non-additive $\Q$, same as in the first experiment, with $\alpha=0.5$
and fix $p=128, s=5$. We measure success not through exact recovery
but through faithful recovery. We say that a trial is a successful if
(1) all relevant variables were recovered and (2) fewer than $20$
variables were marked as relevant overall (true sparsity $s=5$). We
use the same $\lambda$ as before. The probabilities of success are
computed from 40 independent trials and plotted against various values
of $\nu$ in Figure~\ref{Support}(d). Additionally, for $\nu = 0.5$, we
show the number of selected variables versus the sample size as a
box plot in Figure~\ref{Support}(e). As seen, for small to
moderate correlations, AC/DC can successfully recover the relevant
variables with only a small number of false positives.

In the fourth and fifth simulation, we use a softmax function as the ground truth,
\begin{align}
f_0(x_{iS}) = \log \left( \sum_{k=1}^K \exp( \beta_k^\tran x_{iS} ) \right) - \mu.
\end{align}
We generate random unit vectors as $\{\beta_k \in \R^s\}_{k=1,\ldots,K}$ and choose $\mu$ so that $f_0$ has mean-zero. We set $K = 7$ for all the experiments.

\subsubsection{Fourth simulation}
For the \textrm{fourth simulation} (Figure~\ref{Support}f,g), we let
$f_0$ be the softmax function and let $X$ be drawn from the boundary
flat mixture distribution described earlier with the Toeplitz
covariance $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0.5$. We set $s=5$ and
vary $p=128,256,512$. We use the same faithful recovery criteria as in
the third simulation and plot the probability of faithful recovery
against the number of samples in Figure~\ref{Support}(f). The
probabilities are computed over 30 independent trials. Also, for
$p=256$, we show the number of selected variables versus the sample
size as a box plot in Figure~\ref{Support}(g). The softmax function is
more challenging to estimate than the quadratic function; the softmax
function requires about $n > 1{,}500$ samples to achieve the same success
probability as the quadratic function with $n=1{,}000$. Regardless, we
see that increasing the ambient dimension $p$ does not significantly
affect the recovery probability.

@@

\subsubsection{Fifth simulation}
For the \textrm{fifth simulation} (Figure~\ref{fig:ac_v_dc}), we
compare the variables selected via the AC stage and the variables
selected via the DC stage. We use the softmax regression function and
$X$ drawn from the boundary flat mixture distribution with a Toeplitz
covariance and correlation level $\nu = -0.7$. We set $s=5$, $n=500$,
and $p=128$. We perform 30 independent trials and plot the frequency of
variable selection in Figure~\ref{fig:ac_v_dc}. The true variables are
$X_j$ for $j=5,6,7,8,9,10$. We plot the frequency of selection among
only the first 20 variables, that is, $X_j$ for $j=1,\ldots,20$. We do
not plot selection frequencies for variables 21 to 128 because they
are almost never selected by either the AC or DC stage. As can be
seen, the DC stage is slightly helpful in recovering the true
variables but its effect is not significant. We thus believe that the
DC stage, though important in theory, is not as important in practice;
it may be omitted without significant detriment to the overall result.

@@

\begin{figure*}
\includegraphics[width=0.5\textwidth]{figs/ACvDC}
\caption{Frequency of variable selection among the first 20 variables ($X_j$ for $j=1,\ldots,20$) in the AC stage vs. in the DC stage. The true variables are [5,6,7,8,9,10].}
\label{fig:ac_v_dc}
\end{figure*}

@@

We next use the Boston housing data rather than simulated data. This
data set contains 13 covariates, 506 samples and one response variable
indicating housing values in suburbs of Boston. The data and a
detailed description can be found on the UCI Machine Learning
Repository
website\footnote{\url{http://archive.ics.uci.edu/ml/datasets/Housing}}.

@@

Next, in order to quantitatively study the predictive performance, we
run 5-fold cross validation three times, following the same procedure
described above---training, variable selection and refitting.  A plot
of the mean and standard deviation of the predictive mean squared
error (MSE) is shown in Figure \ref{Boston}(c). Since for AC/DC the same
regularization level $\lambda^{(t)}$ may lead to a slightly different number of selected
variables in different folds and runs, the values on the $x$-axis
for AC/DC are not necessarily integers. The figure clearly shows that AC/DC has a 
lower predictive MSE than LASSO.  We also compared the performance of
AC/DC with that of Additive Forward Regression (AFR) presented
in~\cite{Xi:09}, and found that they are similar.  The main advantages
of AC/DC compared with AFR and SpAM are that there are no smoothing
parameters required, and the optimization is formulated
as a convex program, guaranteeing a global optimum.
